ABM simulating short-term mating behaviors

### 2022-06-03 submission/manuscript.Rmd
edited the manuscript based on feedback, close #20

However, it is important to note that although heterosexual men’s total number of short-term mating is equal to that of women, this does not necessarily mean that the total number of men and women who have ever had short-term mating must be equal. Since women have higher standards for short-term mates than men [@buss_sexual_1993], it is possible that the proportion of individuals who meet a potential partner's standards is lower among men than among women, assuming both men and women's desirability values follow the same distribution. Consequently, there may be a smaller proportion of men (vs. the proportion of women) who participate in the mating pool and contribute to the total number of short-term mating. For example, suppose that there is a population of 20 individuals with an equal sex ratio. There are 8 men and 8 women with a desirability value greater than x~1~ and 3 men and 3 women with a value greater than x~2~ (x~1~ < x~2~). Suppose that x~1~ and x~2~ happen to be men and women's standards for short-term mates, then in a long enough period of time, the 8 women who meet men's standards and the 3 men who meet women's standards will have had short-term mating. Since the total number of short-term mating behaviors is the same between men and women, we would expect that in a population of individuals who had any short-term mating, the average number of short-term mating experiences and mates was higher among men than women (e.g., in the aforementioned example, the number is n/3 for men and n/8 for women, where n can be the total number of short-term mating experiences or the number of short-term mates).

@@

We also modeled forming long-term relationships as a background process. A pair of individuals had a possibility to commit to a long-term relationship upon meeting each other. We modeled forming long-term relationships as a behavioral outcome and ignored the psychological processes behind it, although we recognize gender differences may also exist in long-term mating strategy [@buss_sexual_1993; @buss2019mate].

@@

We formulated the following hypotheses in the present study: assuming gender differences in preferences for short-term mating, 1) there would be no gender differences in short-term mating behaviors among all heterosexual men and women, 2) a smaller proportion of heterosexual men had ever engaged into short-term mating compared to the proportion of heterosexual women, 3) in a population of heterosexual individuals who had any short-term mating, men had higher average number of short-term mating behaviors than women, 4) there would be gender differences in short-term mating behaviors among gay men and lesbian women, with gay men being engaged in more such behaviors, and 5) gay men would engage in more short-term mating behaviors compared to heterosexual men.

@@

The present study used agent-based modeling to investigate whether men and women’s differential preferences for short-term mating would result in gender differences in short-term mating behaviors, and if so, then under what conditions they would result in differences in the number of short-term mating experiences and short-term mates. We formulated two hypotheses: 1) gay men would engage in more short-term mating behaviors as compared to lesbian women, and 2) gay men would engage in more short-term mating behaviors as compared to heterosexual men. The results from 1,000 time steps in a model simulating men and women’s mating behaviors provided strong evidence in favor of our hypotheses. First of all, as compared to lesbian women, gay men had higher average numbers of short-term mating experiences and short-term mates. Secondly, gay men also had higher average numbers of short-term mating experiences and short-term mates as compared to heterosexual men. In contrast, we found no gender differences in short-term mating behaviors between heterosexual men and women, although heterosexual men in the mating pool had higher average numbers of short-term mating experiences and short-term mates.

As expected, heterosexual men and women did not differ in short-term mating behaviors despite their differential preferences for short-term mating. This was because heterosexual men and women had an equal total number of both short-term mating experiences and short-term mates. This is not surprising since the sex ratio was 1:1 in our model. It therefore follows that the average number of short-term mating experiences and short-term mates must be equal between men and women as well. However, we did find that when we only looked at individuals in the mating pool (from which more men than women were excluded), men engaged in more short-term mating behaviors as compared to women. This was because there were more women than men in the mating pool, resulting in lower averages among heterosexual women despite the equal numbers of experiences and mates in total.

@@

Using agent-based modeling to simulate short-term mating behaviors, we found when men (vs. women) had a stronger preferences for short-term mating, heterosexual men and women engaged in short-term mating behaviors to the same extent, while gay men engaged in more short-term mating behaviors as compared to lesbian women. Gay men engaged in more short-term mating behaviors than heterosexual men did.

These results highlight the distinction between preferences and behaviors in human mating. Individuals’ mating behaviors do not only depend on one’s own preferences, but are also constrained by partners’ preferences. Future research in human mating should not only focus on the psychological aspect but also pay attention to the interaction between individuals’ psychology and its context. These results also cast doubt to the prevalent belief in the gender differences in short-term mating behaviors, especially among heterosexual individuals. Our findings suggest that there may be factors in the observation process, such as sampling bias, that contribute to the observed differences. Future research in human sexuality should note such possibilities and interpret any observed gender differences in short-term mating behaviors cautiously.

@@

>>>>

However, it is important to note that although heterosexual men’s total number of short-term mating is equal to that of women, this does not necessarily mean that the total number of men and women who have ever had short-term mating must be equal. Empirical studies have found that the proportion of individuals who are sexual inexperienced is slightly higher among men than women in most age groups above 18 years old [@haydon2014prevalence; @ghaznavi2019trends]. As a postulation, there may be a smaller proportion of men (vs. the proportion of women) who participate in the mating pool and contribute to the total number of short-term mating. For example, suppose that there is a population of 20 individuals with an equal sex ratio. Heterosexual short-term mating has happened \(n\) times in this population, but only 3 men and 8 women have had any short-term mating experiences. Since the total number of short-term mating behaviors is the same between men and women, we would expect that in a population of individuals who had any short-term mating, the average number of short-term mating experiences and mates was higher among men than women (e.g., in the aforementioned example, the average number of short-term experiences is \(\frac{n}{3}\) for men and \(\frac{n}{8}\) for women who are in the mating pool).

@@

We also modeled forming long-term relationships as a background process. A pair of individuals had a possibility to commit to a long-term relationship upon meeting each other. This process was modeled at a behavioral level and the psychological processes behind it was ignored, although we recognized gender differences may also exist in long-term mating strategy [@buss_sexual_1993; @buss2019mate].

@@

We formulated the following hypotheses in the present study: assuming gender differences in preferences for short-term mating, 1) there would be no gender differences in short-term mating behaviors between heterosexual men and women when the full population was considered, 2) a smaller proportion of heterosexual men (vs. heterosexual women) had ever engaged in short-term mating, 3) in a population of heterosexual individuals who had any short-term mating, men had higher average number of short-term mating behaviors than women, 4) gay men would engage in more short-term mating behaviors compared to lesbian women, and 5) gay men would engage in more short-term mating behaviors compared to heterosexual men.

@@

The present study used agent-based modeling to investigate whether men and women’s differential preferences for short-term mating would result in gender differences in short-term mating behaviors, and if so, then under what conditions they would result in differences in the number of short-term mating experiences and short-term mates. We formulated five hypotheses: 1) there would be no gender differences in short-term mating behaviors between heterosexual men and women when the full population was considered, 2) a smaller proportion of heterosexual men (vs. heterosexual women) had ever engaged in short-term mating, 3) in a population of heterosexual individuals who had any short-term mating, men had higher average number of short-term mating behaviors than women, 4) gay men would engage in more short-term mating behaviors compared to lesbian women, and 5) gay men would engage in more short-term mating behaviors compared to heterosexual men. The results from 1,000 time steps in a model simulating men and women’s mating behaviors provided strong evidence in favor of our hypotheses. First of all, we found no gender differences in short-term mating behaviors between heterosexual men and women when the full population was considered. However, compared to heterosexual women, there was a smaller proportion of heterosexual men participating in the mating pool, but they had higher average numbers of short-term mating experiences and short-term mates. Secondly, as compared to lesbian women, gay men had higher average numbers of short-term mating experiences and short-term mates. Finally, gay men also had higher average numbers of short-term mating experiences and short-term mates as compared to heterosexual men.

As expected, when the full population was considered, heterosexual men and women did not differ in short-term mating behaviors despite their differential preferences for short-term mating. This was because heterosexual men and women had an equal total number of both short-term mating experiences and short-term mates. This is not surprising since the sex ratio was 1:1 in our model. It therefore follows that the average number of short-term mating experiences and short-term mates must be equal between men and women as well. However, we did find that when we only looked at individuals in the mating pool (from which more men than women were excluded), men engaged in more short-term mating behaviors as compared to women. This was because there were more women than men in the mating pool, resulting in lower averages among heterosexual women despite the equal numbers of experiences and mates in total.

@@

Using agent-based modeling to simulate short-term mating behaviors, we found when men (vs. women) had a stronger preferences for short-term mating, heterosexual men and women engaged in short-term mating behaviors to the same extent when the full population was considered, while men in the mating pool had more such behaviors than women in the mating pool. Gay men engaged in more short-term mating behaviors compared to both lesbian women and heterosexual men.

These results highlight the distinction between preferences and behaviors in human mating. Individuals’ mating behaviors do not only depend on one’s own preferences, but are also constrained by partners’ preferences. Future research in human mating should not only focus on the psychological aspect but also pay attention to the interaction between individuals’ psychology and its behavioral context. These results also cast doubt to the prevalent belief in the gender differences in short-term mating behaviors, especially among heterosexual individuals. Our findings suggest that there may be factors in the observation process, such as sampling bias, that contribute to the observed differences. Future research in human sexuality should note such possibilities and interpret any observed gender differences in short-term mating behaviors with caution.

@@

We thank Gu Li and Jiannan Shi for providing helpful feedback in the development of this study.

### 2022-06-03 submission/manuscript.tex
edited the manuscript based on feedback, close #20

However, it is important to note that although heterosexual men's total
number of short-term mating is equal to that of women, this does not
necessarily mean that the total number of men and women who have ever
had short-term mating must be equal. Since women have higher standards
for short-term mates than men (Buss \& Schmitt, 1993), it is possible
that the proportion of individuals who meet a potential partner's
standards is lower among men than among women, assuming both men and
women's desirability values follow the same distribution. Consequently,
there may be a smaller proportion of men (vs.~the proportion of women)
who participate in the mating pool and contribute to the total number of
short-term mating. For example, suppose that there is a population of 20
individuals with an equal sex ratio. There are 8 men and 8 women with a
desirability value greater than x\textsubscript{1} and 3 men and 3 women
with a value greater than x\textsubscript{2} (x\textsubscript{1}
\textless{} x\textsubscript{2}). Suppose that x\textsubscript{1} and
x\textsubscript{2} happen to be men and women's standards for short-term
mates, then in a long enough period of time, the 8 women who meet men's
standards and the 3 men who meet women's standards will have had
short-term mating. Since the total number of short-term mating behaviors
is the same between men and women, we would expect that in a population
of individuals who had any short-term mating, the average number of
short-term mating experiences and mates was higher among men than women
(e.g., in the aforementioned example, the number is n/3 for men and n/8
for women, where n can be the total number of short-term mating
experiences or the number of short-term mates).

@@

We also modeled forming long-term relationships as a background process.
A pair of individuals had a possibility to commit to a long-term
relationship upon meeting each other. We modeled forming long-term
relationships as a behavioral outcome and ignored the psychological
processes behind it, although we recognize gender differences may also
exist in long-term mating strategy (Buss \& Schmitt, 1993; Buss \&
Schmitt, 2019).

@@

We formulated the following hypotheses in the present study: assuming
gender differences in preferences for short-term mating, 1) there would
be no gender differences in short-term mating behaviors among all
heterosexual men and women, 2) a smaller proportion of heterosexual men
had ever engaged into short-term mating compared to the proportion of
heterosexual women, 3) in a population of heterosexual individuals who
had any short-term mating, men had higher average number of short-term
mating behaviors than women, 4) there would be gender differences in
short-term mating behaviors among gay men and lesbian women, with gay
men being engaged in more such behaviors, and 5) gay men would engage in
more short-term mating behaviors compared to heterosexual men.

@@

The present study used agent-based modeling to investigate whether men
and women's differential preferences for short-term mating would result
in gender differences in short-term mating behaviors, and if so, then
under what conditions they would result in differences in the number of
short-term mating experiences and short-term mates. We formulated two
hypotheses: 1) gay men would engage in more short-term mating behaviors
as compared to lesbian women, and 2) gay men would engage in more
short-term mating behaviors as compared to heterosexual men. The results
from 1,000 time steps in a model simulating men and women's mating
behaviors provided strong evidence in favor of our hypotheses. First of
all, as compared to lesbian women, gay men had higher average numbers of
short-term mating experiences and short-term mates. Secondly, gay men
also had higher average numbers of short-term mating experiences and
short-term mates as compared to heterosexual men. In contrast, we found
no gender differences in short-term mating behaviors between
heterosexual men and women, although heterosexual men in the mating pool
had higher average numbers of short-term mating experiences and
short-term mates.

As expected, heterosexual men and women did not differ in short-term
mating behaviors despite their differential preferences for short-term
mating. This was because heterosexual men and women had an equal total
number of both short-term mating experiences and short-term mates. This
is not surprising since the sex ratio was 1:1 in our model. It therefore
follows that the average number of short-term mating experiences and
short-term mates must be equal between men and women as well. However,
we did find that when we only looked at individuals in the mating pool
(from which more men than women were excluded), men engaged in more
short-term mating behaviors as compared to women. This was because there
were more women than men in the mating pool, resulting in lower averages
among heterosexual women despite the equal numbers of experiences and
mates in total.

@@

Using agent-based modeling to simulate short-term mating behaviors, we
found when men (vs.~women) had a stronger preferences for short-term
mating, heterosexual men and women engaged in short-term mating
behaviors to the same extent, while gay men engaged in more short-term
mating behaviors as compared to lesbian women. Gay men engaged in more
short-term mating behaviors than heterosexual men did.

These results highlight the distinction between preferences and
behaviors in human mating. Individuals' mating behaviors do not only
depend on one's own preferences, but are also constrained by partners'
preferences. Future research in human mating should not only focus on
the psychological aspect but also pay attention to the interaction
between individuals' psychology and its context. These results also cast
doubt to the prevalent belief in the gender differences in short-term
mating behaviors, especially among heterosexual individuals. Our
findings suggest that there may be factors in the observation process,
such as sampling bias, that contribute to the observed differences.
Future research in human sexuality should note such possibilities and
interpret any observed gender differences in short-term mating behaviors
cautiously.

@@

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-wilensky_1999}{}}%
Wilensky, U. (1999). \emph{NetLogo}
{[}Http://ccl.northwestern.edu/netlogo/{]}. Center for Connected
Learning; Computer-Based Modeling.
\url{http://ccl.northwestern.edu/netlogo/}

>>>>

However, it is important to note that although heterosexual men's total
number of short-term mating is equal to that of women, this does not
necessarily mean that the total number of men and women who have ever
had short-term mating must be equal. Empirical studies have found that
the proportion of individuals who are sexual inexperienced is slightly
higher among men than women in most age groups above 18 years old
(Ghaznavi et al., 2019; Haydon et al., 2014). As a postulation, there
may be a smaller proportion of men (vs.~the proportion of women) who
participate in the mating pool and contribute to the total number of
short-term mating. For example, suppose that there is a population of 20
individuals with an equal sex ratio. Heterosexual short-term mating has
happened \(n\) times in this population, but only 3 men and 8 women have
had any short-term mating experiences. Since the total number of
short-term mating behaviors is the same between men and women, we would
expect that in a population of individuals who had any short-term
mating, the average number of short-term mating experiences and mates
was higher among men than women (e.g., in the aforementioned example,
the average number of short-term experiences is \(\frac{n}{3}\) for men
and \(\frac{n}{8}\) for women who are in the mating pool).

@@

We also modeled forming long-term relationships as a background process.
A pair of individuals had a possibility to commit to a long-term
relationship upon meeting each other. This process was modeled at a
behavioral level and the psychological processes behind it was ignored,
although we recognized gender differences may also exist in long-term
mating strategy (Buss \& Schmitt, 1993; Buss \& Schmitt, 2019).

@@

We formulated the following hypotheses in the present study: assuming
gender differences in preferences for short-term mating, 1) there would
be no gender differences in short-term mating behaviors between
heterosexual men and women when the full population was considered, 2) a
smaller proportion of heterosexual men (vs.~heterosexual women) had ever
engaged in short-term mating, 3) in a population of heterosexual
individuals who had any short-term mating, men had higher average number
of short-term mating behaviors than women, 4) gay men would engage in
more short-term mating behaviors compared to lesbian women, and 5) gay
men would engage in more short-term mating behaviors compared to
heterosexual men.

@@

The present study used agent-based modeling to investigate whether men
and women's differential preferences for short-term mating would result
in gender differences in short-term mating behaviors, and if so, then
under what conditions they would result in differences in the number of
short-term mating experiences and short-term mates. We formulated five
hypotheses: 1) there would be no gender differences in short-term mating
behaviors between heterosexual men and women when the full population
was considered, 2) a smaller proportion of heterosexual men
(vs.~heterosexual women) had ever engaged in short-term mating, 3) in a
population of heterosexual individuals who had any short-term mating,
men had higher average number of short-term mating behaviors than women,
4) gay men would engage in more short-term mating behaviors compared to
lesbian women, and 5) gay men would engage in more short-term mating
behaviors compared to heterosexual men. The results from 1,000 time
steps in a model simulating men and women's mating behaviors provided
strong evidence in favor of our hypotheses. First of all, we found no
gender differences in short-term mating behaviors between heterosexual
men and women when the full population was considered. However, compared
to heterosexual women, there was a smaller proportion of heterosexual
men participating in the mating pool, but they had higher average
numbers of short-term mating experiences and short-term mates. Secondly,
as compared to lesbian women, gay men had higher average numbers of
short-term mating experiences and short-term mates. Finally, gay men
also had higher average numbers of short-term mating experiences and
short-term mates as compared to heterosexual men.

As expected, when the full population was considered, heterosexual men
and women did not differ in short-term mating behaviors despite their
differential preferences for short-term mating. This was because
heterosexual men and women had an equal total number of both short-term
mating experiences and short-term mates. This is not surprising since
the sex ratio was 1:1 in our model. It therefore follows that the
average number of short-term mating experiences and short-term mates
must be equal between men and women as well. However, we did find that
when we only looked at individuals in the mating pool (from which more
men than women were excluded), men engaged in more short-term mating
behaviors as compared to women. This was because there were more women
than men in the mating pool, resulting in lower averages among
heterosexual women despite the equal numbers of experiences and mates in
total.

@@

Using agent-based modeling to simulate short-term mating behaviors, we
found when men (vs.~women) had a stronger preferences for short-term
mating, heterosexual men and women engaged in short-term mating
behaviors to the same extent when the full population was considered,
while men in the mating pool had more such behaviors than women in the
mating pool. Gay men engaged in more short-term mating behaviors
compared to both lesbian women and heterosexual men.

These results highlight the distinction between preferences and
behaviors in human mating. Individuals' mating behaviors do not only
depend on one's own preferences, but are also constrained by partners'
preferences. Future research in human mating should not only focus on
the psychological aspect but also pay attention to the interaction
between individuals' psychology and its behavioral context. These
results also cast doubt to the prevalent belief in the gender
differences in short-term mating behaviors, especially among
heterosexual individuals. Our findings suggest that there may be factors
in the observation process, such as sampling bias, that contribute to
the observed differences. Future research in human sexuality should note
such possibilities and interpret any observed gender differences in
short-term mating behaviors with caution.

@@

\leavevmode\vadjust pre{\hypertarget{ref-ghaznavi2019trends}{}}%
Ghaznavi, C., Sakamoto, H., Yoneoka, D., Nomura, S., Shibuya, K., \&
Ueda, P. (2019). Trends in heterosexual inexperience among young adults
in japan: Analysis of national surveys, 1987--2015. \emph{BMC Public
Health}, \emph{19}(1), 1--10.

@@

\leavevmode\vadjust pre{\hypertarget{ref-haydon2014prevalence}{}}%
Haydon, A. A., Cheng, M. M., Herring, A. H., McRee, A.-L., \& Halpern,
C. T. (2014). Prevalence and predictors of sexual inexperience in
adulthood. \emph{Archives of Sexual Behavior}, \emph{43}(2), 221--230.

@@

\leavevmode\vadjust pre{\hypertarget{ref-wilensky_1999}{}}%
Wilensky, U. (1999). \emph{NetLogo}. Center for Connected Learning;
Computer-Based Modeling. \url{http://ccl.northwestern.edu/netlogo/}
